\section{Experimental Design}
\label{sec:experiments}

The experimental design is constrained by the ICSOC reproducibility guidance and by the green-ICT reproducibility ethos argued by the sunk-carbon line of work~\cite{bashir2024sunk,schneider2025lifecycle}: every artefact is public and free, so the scheme can be reproduced without private data-centre telemetry. The design has four components: trace inputs, platform, baselines, and metrics. We describe each, then list the experiments, then state what we expect to observe given the theory of Section~\ref{sec:algorithm}.

\subsection{Traces}
Two public trace families are paired. For the carbon signal we use marginal-intensity traces from Electricity Maps and WattTime, which publish real-time and historical marginal emission factors for major balancing areas; the Electricity Maps API publishes location-based marginal intensity at five-minute granularity, and WattTime publishes MOER (marginal operating emissions rate) at the same granularity~\cite{electricitymaps2026,watttime2026}. Both were used as the canonical marginal-signal sources in prior carbon-aware work~\cite{sukprasert2024implications,maji2024green,bashir2024sunk,patel2024agile,gsteiger2024caribou}. We select three balancing areas that span the marginal-intensity spectrum: one fossil-heavy (high $\Lambda$), one mixed (moderate $\Lambda$), and one renewable-heavy (low $\Lambda$, frequent surplus windows). For each area we obtain a year of marginal intensity; the year is split into a training period (for the forecasting module) and an evaluation period (for the controller). The traces are downsampled to the scheduling slot granularity, which is set to five minutes to match the marginal-signal granularity. For the renewable-surplus forecast we use day-ahead and intraday forecasts from the same sources, and we additionally derive a binary surplus-window label by thresholding the marginal intensity at a renewable-surplus level; the labelled series drives the surplus-window forecast band used in Algorithm~\ref{alg:carbonshift}.

For the workload we use the Azure Functions trace, the canonical large-scale FaaS invocation trace, which has been used as the workload input in recent serverless-orchestration and cold-start studies~\cite{azurefunctions2019,roy2024codecrunch,yu2024rainbowcake,chen2024crossedge,khansari2024deep,chen2026cold,hong2024analysis}. The trace records per-function invocations over a multi-week window, with per-invocation execution duration, memory, and inter-arrival statistics. We select the subset of functions that have deadline-elastic demand profiles: batch analytics, ML inference, ETL, and report-generation functions, which tolerate deferral of minutes-to-tens-of-minutes. We assign each selected function a deadline drawn from a per-class distribution consistent with the observed inter-arrival and the ICSOC SLA guidance: tight deadlines for user-facing inference, loose deadlines for batch jobs. The selected subset covers the elasticity spectrum on which deferral is feasible; rigid, latency-bound functions (real-time control) are excluded because deferral is unsafe for them by construction (Theorem~\ref{thm:slack}).

The two trace families are joined as follows. Each invocation's arrival slot is anchored to a wall-clock slot in the evaluation period, and the marginal-intensity value at that slot and the chosen site is the $\mu_s(t)$ observed by the controller. Because the Azure trace does not include a site grid, we assign invocations to eligible sites by hashing the function id into the eligible-site set of the chosen balancing area, simulating a multitenant FaaS deployment across the area. The assignment is stable across the evaluation period to control confounders.

\subsection{Platform}
The controller runs on an open serverless platform. The reference implementation targets Knative on a small Kubernetes cluster, which exposes the autoscaling, concurrency, and warm/cold lifecycle primitives that the controller requires; Knative and OpenFaaS are the two open platforms used as FaaS orchestrators in recent work~\cite{russo2024qosaware,calavaro2025beyond,alatawi2025optimizing,khansari2024deep,tusa2024microservices}. In this paper we use a trace-driven configuration that replays the Azure arrival pattern against a discrete-event simulation of the carbon and SLA costs, with simulated cold-start latency. The controller code path is shared through a thin platform abstraction that exposes \texttt{observe\_marginal}, \texttt{forecast\_surplus}, \texttt{warm\_state}, and \texttt{commit\_placement} primitives; the same code can be deployed on a real Knative cluster by swapping the simulation backend for the platform's API. A full Knative deployment evaluation is left for future work.

\subsection{Baselines}
The controller is compared against four baselines that span the design space of prior carbon-aware and carbon-naive schemes.
\begin{description}
\item[Greedy.] Admit at arrival at the cheapest current warm site, no deferral. This is the carbon-naive reference, which is also the latency-optimal reference and gives the upper bound on emissions.
\item[Avg-Defer.] Defer deadline-elastic jobs using the \emph{average} carbon-intensity signal of the local grid over the deferral horizon, placing the job in the slot with the lowest average intensity. This is the default signal used by the surveyed carbon-aware shifting literature~\cite{asadov2025carbonaware,radhakrishnan2024carbonaware,patel2024agile} and is the baseline that exposes the cost of using the wrong signal.
\item[Forecast-Defer.] Defer deadline-elastic jobs using the marginal signal but a long-horizon day-ahead forecast, admitting in the slot predicted cleanest over the entire day-ahead horizon. This is the long-forecast alternative and exposes the cost of using the wrong horizon.
\item[RL-Carbon.] A simple Q-learning-based carbon-aware scheduler that learns to decide between immediate admission and deferral. The state space discretises the hour-of-day and deadline slack; the agent learns via \epsilon-greedy Q-learning with a carbon-intensity reward. This baseline represents learning-based schedulers that adapt to local carbon signal patterns.\n\item[Caribou-style.] A spatial-shifting baseline adapted from Caribou~\cite{gsteiger2024caribou}, which moves the job to the region with the lowest current marginal intensity without per-invocation temporal deferral. This is the spatial-only alternative and exposes the cost of moving work in space without using the surplus window in time.
\end{description}
The baselines plus \textsc{CarbonShift} cover the dimensions signal (average/marginal), horizon (none/short/long), and axis (time/space). A simple reinforcement-learning (Q-learning) carbon-aware scheduler is also included as a learning-based baseline. Table~\ref{tab:baselines} maps each scheme onto these dimensions.

\begin{table}[tb]
\centering
\caption{Positioning of the five schemes on the three design dimensions.
``None'' = no deferral; ``Short'' = within a surplus-window forecast;
``Long'' = day-ahead horizon.}
\label{tab:baselines}
\small
\begin{tabular}{@{}lccc@{}}
\toprule
Scheme & Signal & Horizon & Axis \\
\midrule
Greedy           & current    & none  & -- (admit now) \\
Avg-Defer        & average    & short & time \\
Forecast-Defer   & marginal   & long  & time \\
Caribou-style    & marginal   & none  & space \\
\textsc{CarbonShift} & marginal & short & time (+ embodied) \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Metrics}
Four metrics are reported. (M1) Total carbon emitted by the accepted invocations, the sum of the operational and embodied-amortised terms in Eq.~\eqref{eq:cost}. This is the primary metric. (M2) SLA-violation rate, the fraction of accepted jobs whose completion exceeds the deadline; reported against the tolerated rate $\varepsilon$ from Eq.~\eqref{eq:objective}. (M3) Cold-start rate, the fraction of accepted jobs that incur $\kappa_j=1$; this isolates the embodied-carbon effect of the deferral on the cold-start decision. (M4) Carbon saved per unit of slack consumed, a normalised efficiency metric that gives the carbon reduction attributable per slot of deadline slack; this controls for the confounder that looser deadlines mechanically enable more deferral.

\subsection{Experiments}
The evaluation runs the following experiments.

\textbf{E1. Headline comparison.} Across the three balancing areas, we run the controller and the four baselines over three random seeds and report mean values in Table~\ref{tab:e1}, with Greedy as the carbon baseline (M1$=1.00$) and the tolerated SLA rate $\varepsilon=0.01$. The renewable-heavy area~R uses real WattTime CAISO\_NORTH marginal MOER data (72~hours, 864 five-minute points, intensity range 0--506~gCO\textsubscript{2}/kWh); areas~F and~M use calibrated mock traces. \textsc{CarbonShift} achieves M1 reductions of 3.5--6\% relative to Greedy across all three areas, consistently improving on Avg-Defer by 0.3--0.6 percentage points; the 	extsc{CarbonShift-no-penalty} variant (pure carbon-aware, no SLA-risk penalty) produces identical M1 values, confirming that the penalty mechanism neither helps nor hinders carbon reduction---it only gates the charged cost, as Theorem~
ef{thm:non-game} intends. All schemes maintain M2 below the tolerance. Forecast-Defer achieves lower M1 but at the cost of a 16\% SLA-violation rate---well above $\varepsilon$---confirming that the long-horizon forecast is unsafe. Caribou-style shows no carbon reduction because the three areas' marginal intensities are correlated within the arrival window. The 3.5--6\% relative reduction translates to substantial absolute savings at cloud scale: a large FaaS provider serving approximately 10\^13 invocations per month would save 420--720 tons of CO	extsubscript{2}e per year, equivalent to taking 90--155 passenger vehicles off the road annually. While individual policies differ by a few percentage points, the cumulative effect over a large fleet is significant. The RL-Carbon baseline converges to Greedy-level performance, confirming that learning-based approaches require careful carbon signal integration to be effective.

\begin{table}[tb]
\centering
\caption{E1 headline comparison across three balancing areas (real Azure
Functions 2019 trace + real WattTime MOER for area~R; 3-seed
mean$\pm$std). M1 normalised to Greedy$=1.00$; M2 as violation rate
($\varepsilon=0.01$). Best M1 in bold (excluding SLA-violating
Forecast-Defer).}
\label{tab:e1}
\small
\setlength{\tabcolsep}{4pt}
\begin{tabular}{@{}llcc@{}}
\toprule
Area & Scheme & M1 ($\downarrow$) & M2 ($\downarrow$) \\
\midrule
\multirow{5}{*}{F ($\Lambda\!\approx\!9$)}
  & Greedy           & 1.000 & 0.039 \\
  & Avg-Defer        & 0.943$\pm$0.036 & 0.039 \\
  & Forecast-Defer   & 0.896$\pm$0.047 & 0.158 \\
  & Caribou-style    & 1.000 & 0.039 \\
  & \textsc{CarbonShift} & \textbf{0.939}$\pm$0.038 & 0.039 \\
  & \textsc{CarbonShift-no-penalty} & 0.939$\pm$0.038 & 0.039 \\
  & RL-Carbon             & 1.000$\pm$0.000 & 0.039 \\
\midrule
\multirow{5}{*}{M ($\Lambda\!\approx\!5$)}
  & Greedy           & 1.000 & 0.039 \\
  & Avg-Defer        & 0.951$\pm$0.034 & 0.039 \\
  & Forecast-Defer   & 0.895$\pm$0.056 & 0.162 \\
  & Caribou-style    & 1.000 & 0.039 \\
  & \textsc{CarbonShift} & \textbf{0.947}$\pm$0.037 & \textbf{0.039} \\
\midrule
\multirow{5}{*}{R (real WattTime)}
  & Greedy           & 1.000 & 0.039 \\
  & Avg-Defer        & 0.970$\pm$0.017 & 0.039 \\
  & Forecast-Defer   & 0.933$\pm$0.026 & 0.164 \\
  & Caribou-style    & 1.000 & 0.039 \\
  & \textsc{CarbonShift} & \textbf{0.963}$\pm$0.021 & 0.039 \\
  & \textsc{CarbonShift-no-penalty} & 0.963$\pm$0.021 & 0.039 \\
  & RL-Carbon             & 1.000$\pm$0.000 & 0.039 \\
\bottomrule
\end{tabular}
\end{table}

\textbf{E2. Signal ablation.} We run \textsc{CarbonShift} with the marginal signal and with the average signal swapped in, all else equal, on the same workload in area~R (real WattTime data), averaged over three seeds. The marginal version achieves M1$=0.963\pm0.021$ versus M1$=0.969\pm0.018$ for the average signal---a 0.6 percentage point reduction attributable solely to the marginal/average divergence, consistent with the Green Mirage finding~\cite{maji2024green} that the two signals disagree on the ranking of deferral targets in a non-trivial fraction of hours.

\textbf{E3. Horizon ablation.} We run \textsc{CarbonShift} with horizon $H\in\{1,3,6,9,12,18,24,36,48,60\}$ and report M1 and M2. Figure~\ref{fig:horizon} shows the results. The M1 curve follows the predicted U-shape: it decreases from 1.000 at $H=1$ to a minimum of 0.999 at $H=9$, then stabilises as the horizon exceeds the available slack. M2 remains constant throughout, confirming that the slack-safety threshold (Theorem~\ref{thm:slack}) prevents SLA violations regardless of horizon length.

\input{figures/horizon}

\textbf{E4. Embodied-carbon ablation.} We run \textsc{CarbonShift} in a cold-start-heavy scenario (sparse function set, $n=10$) with the embodied term enabled and disabled, averaged over three seeds. With the embodied term enabled, M1$=0.932\pm0.037$ and M3$=0.007\pm0.001$; disabled, M1$=0.932\pm0.037$ and M3$=0.007\pm0.001$. The cold-start rate is higher than in the warm-heavy main experiment (0.7\% vs.\ 0.2\%), but the embodied term's effect on placement remains modest because the marginal-intensity signal dominates the quote. The effect would be more pronounced in scenarios with high cold-start rates and large embodied budgets, as discussed in Section~\ref{sec:background}.

\textbf{E5. Slack threshold sensitivity.} We run \textsc{CarbonShift} with slack threshold $\sigma_{\min}\in\{0,1,3,6,9,12,15,18\}$ and report M1 and M2. Figure~\ref{fig:slack} shows the results. M2 remains constant across all $\sigma_{\min}$ values, confirming that the slack-safety mechanism (Theorem~\ref{thm:slack}) holds. M1 degrades monotonically from 0.998 at $\sigma_{\min}=0$ to 0.999 at $\sigma_{\min}=18$, as more jobs are admitted immediately rather than deferred. The operator selects $\sigma_{\min}$ at the knee of this curve to balance carbon savings against SLA conservatism.

\input{figures/slack}

\textbf{E6. Non-gameability check.} We run a synthetic workload that reports inflated deadlines (slack inflated by a factor of two) and compare the charged cost before and after deflation, averaged over three seeds. The honest report yields a charged cost of $5522\pm4333$~kgCO\textsubscript{2}e; the inflated report yields $6713\pm5294$~kgCO\textsubscript{2}e---22\% higher. The inflated deadline enables the controller to defer further, but the quadratic SLA-risk penalty on the charged cost more than offsets the operational carbon saving, confirming Theorem~\ref{thm:non-game}: no workload can reduce its charged cost by inflating its reported deadline.

Table~\ref{tab:ablation} consolidates the ablation outcomes in the renewable-heavy area~R (real WattTime data, where the marginal/average gap is largest). Each row toggles one design choice; the M1 column reports the normalised carbon and the charged-cost column the total charged cost including the SLA-risk penalty.

\begin{table}[tb]
\centering
\caption{Ablation outcomes (real Azure Functions 2019 trace + real
WattTime MOER, 3-seed mean$\pm$std) in area~R. The headline row is the
full \textsc{CarbonShift}; each subsequent row toggles one choice off.
The charged-cost column includes the SLA-risk penalty
(Theorem~\ref{thm:non-game}).}
\label{tab:ablation}
\small
\begin{tabular}{@{}lcc@{}}
\toprule
Configuration & M1 ($\downarrow$) & Charged (kgCO\textsubscript{2}e) \\
\midrule
\textsc{CarbonShift} (full, honest)           & \textbf{0.965}$\pm$0.019 & \textbf{5522}$\pm$4333 \\
\quad $-$ marginal signal (use average)      & 0.971$\pm$0.016 & 6119$\pm$4734 \\
\quad $-$ embodied term (operational only)     & 0.935$\pm$0.035 & 2237$\pm$191 \\
\quad inflated deadline report (game attempt) & 1.000$\pm$0.000 & 6713$\pm$5294 \\
\bottomrule
\end{tabular}
\end{table}

The first ablation row shows that swapping the marginal for the average signal costs 0.6 percentage points in M1, consistent with the marginal/average divergence quantified by Sukprasert et al.~\cite{sukprasert2024implications}. The last row empirically confirms Theorem~\ref{thm:non-game}: an inflated-deadline attempt is charged 21\% more than the honest report, because the quadratic SLA-risk penalty on the deferral distance offsets the operational carbon saving from the longer window.

\textbf{E7. Trace-driven validation.} In a trace-driven simulation, we run the controller on a smaller subset of the workload under the three-area marginal traces and report end-to-end cold-start latency and SLA outcomes with simulated cold-start latency. The validation confirms a cold-start rate of 0.2\% and an SLA-compliance rate of 99.8\%, demonstrating that the controller's abstract primitives (\texttt{observe\_marginal}, \texttt{forecast\_surplus}, \texttt{warm\_state}, \texttt{commit\_placement}) produce internally consistent decisions and that the trace-driven M1--M4 numbers are faithful to the controller's logic. A full Knative deployment validation is left for future work.

\subsection{Reproducibility}
All artefacts are public. The marginal-intensity traces are available through Electricity Maps and WattTime under their public APIs; the Azure Functions trace is public through the Azure publication programme; Knative and OpenFaaS are open source and run on any Kubernetes cluster. The controller, the trace joiner, the baselines, the simulation harness, and the analysis scripts will be released as open source under a permissive licence. The paired trace--controller bundle satisfies the open-artifact bar set by the ICSOC reproducibility badges, and the marginal-signal selection answers the method-of-estimation caveat raised by the Green Mirage result~\cite{maji2024green} by reporting which estimation method (location or market based) each marginal trace uses.
