\section{Introduction}
\label{sec:intro}

Cloud service systems now consume a fast-growing fraction of global electricity, and the corresponding carbon footprint has made green ICT and sustainability a first-class concern in cloud--edge continuum research~\cite{patel2024modeling,aldulaimy2024computing,gill2024modern}. Function-as-a-Service (FaaS) is the dominant delivery model on this continuum: short-lived, stateless functions are provisioned on demand, billed per invocation, and presented as an easy target for sustainability optimisation because each invocation is individually schedulable~\cite{calavaro2025beyond,russo2024qosaware}. The natural lever that carbon-aware orchestrators pull on such workloads is \emph{deferral}: a deadline-elastic invocation is held back from execution and run later, at a time when the data centre that executes it is powered by a cleaner generation mix. The premise is simple, and it has produced a wave of carbon-aware workload-shifting schemes over the last two years~\cite{asadov2025carbonaware,patel2024agile,radhakrishnan2024carbonaware,gsteiger2024caribou,rodrigues2025carbonaware}.

This paper starts from the observation that two of the assumptions baked into that body of work are, jointly, wrong. The first assumption is the carbon signal. Existing schedulers largely use the \emph{average} carbon intensity of the local grid, the headline number that operators quote and that the dominant public datasets publish. Average intensity records the emissions of the average electron; it does not record the emissions of \emph{the next} electron. When a workload is deferred, the electricity that a data centre consumes because of that workload is, by definition, marginal: it is served by the generator that ramps to meet the incremental demand. Recent measurement work has shown that the choice between average and marginal signals changes the ranking of deferral targets and, in a non-trivial fraction of hours, reverses the recommended action altogether~\cite{sukprasert2024implications,maji2024green}. Worse, deferring a job by using an average signal can move execution into an hour whose marginal intensity is \emph{higher} than the hour the job was deferred from, so the carbon budget goes up rather than down. The sunk-carbon critique sharpens the point further: once infrastructure is built, the carbon already paid cannot be recovered by waiting for a cleaner hour, and optimisations that ignore embodied cost misallocate deferrals~\cite{bashir2024sunk}.

The second assumption is the horizon. Carbon-aware schemes either schedule greedily on the current intensity reading or solve a horizon-constrained optimisation against a long-run renewable-share forecast. Neither matches how renewable surpluses arrive in real grids: as short, sharp, weather-driven windows that follow ramp events and last tens of minutes to a few hours~\cite{singh2025riskaware,chen2024towards}. A greedy scheduler executes a deadline-elastic job during a moderately clean hour and then cannot use the much cleaner window that opens an hour later; a long-forecast scheduler knows a clean window exists on average but not within the job's deadline, so it either defers too aggressively and risks SLA violation or not at all.

We argue that both assumptions should be replaced simultaneously, and that doing so produces a problem that has not been analysed. The scheduler should respond to a real-time \emph{marginal} carbon-intensity signal, should treat an imminent renewable-surplus window as a forecast band whose existence and length are predicted online, and should be charged for the \emph{embodied} carbon of the edge instance it uses, amortising that embodied cost per served request whenever the instance is reused instead of cold-started. The last point closes the loop between deferral, cold starts, and sustainability: a deferral that avoids a dirty hour but triggers a cold start in a clean hour can be net-negative if the cold start's amortised embodied cost exceeds the operational savings.

This paper makes three contributions.

\input{figures/framework_img}

\noindent Figure~\ref{fig:framework} previews the framework. The controller consumes a real-time marginal-intensity signal and a short-horizon surplus forecast, tracks the warm/cold and embodied-cost state of every eligible site, and commits the placement that minimises the carbon quote of Eq.~\eqref{eq:quote}. The dashed warm-state feedback closes the loop between deferral and embodied amortisation. The remainder of the paper formalises this framework and gives its guarantees.

First, we formalise marginal-carbon-aware serverless deferral as an online problem on a cloud--edge continuum in which each arrival is a deadline-elastic FaaS invocation characterised by an execution time, a deadline, an eligible placement set, and a per-site embodied-carbon cost. The objective is the total carbon emitted by serving the accepted invocations, comprising operational carbon charged at the (marginal) intensity of the time and site of execution plus an embodied term amortised per served request. The offline optimum knows the full marginal-intensity trajectory, the full arrival sequence, and all surplus windows; the online algorithm knows none of these and must commit placements and deferral decisions irrevocably.

Second, we give a competitive algorithm, \textsc{CarbonShift}, with provable guarantees. We prove that on adversarial marginal-intensity sequences the algorithm is $\Theta(\Lambda)$-competitive, where $\Lambda$ is the ratio between the highest and lowest marginal intensities that any accepted execution can experience, and that the bound is tight up to a constant. On bounded-variation sequences---sequences whose per-step variation is bounded, which match the autocorrelation observed in real grid traces---we prove the guarantee collapses to a small constant that depends only on the variation budget and the deadline slack, and we give the slack threshold below which deferral must be disabled. We further prove that the marginal signal is non-gameable by deadline inflation: a workload cannot reduce its carbon by reporting a longer deadline, because the cost charged to that workload increases with the slack it consumes.

Third, we design an experimental study that pairs public marginal-intensity traces with public FaaS invocation traces and runs the controller on an open serverless platform. The design evaluates total emissions, SLA compliance, and cold-start cost against four baseline policies spanning greedy, average-intensity-deferral, long-forecast-deferral, and the recent geospatial-shifting baseline. The design is intentionally built on free, public artefacts so that the scheme is reproducible without access to private data-centre telemetry, in line with the ICSOC reproducibility guidance and the open-artifact ethos argued by the sunk-carbon line of work~\cite{bashir2024sunk,schneider2025lifecycle}.

The rest of the paper is organised as follows. Section~\ref{sec:background} reviews the carbon-signal, continuum, and serverless-cold-start background and isolates the gap that the model targets. Section~\ref{sec:model} formalises the problem and the offline optimum. Section~\ref{sec:algorithm} gives the algorithm and the competitive-ratio, slack-safety, and non-gameability theorems. Section~\ref{sec:experiments} gives the experimental design. Section~\ref{sec:related} positions the scheme against the closest prior work, and Section~\ref{sec:conclusion} concludes.
